---
title: "Tidy Tuesday - Scooby Doo dataset"
author: "Amy Mitchell-Whittington"
date: "16/07/2021"
output: github_document
---

## Scooby Doo monsters

This week's dataset comes from [Kaggle](https://www.kaggle.com/williamschooleman/scoobydoo-complete), manually aggregated by [plummye](https://www.kaggle.com/williamschooleman).  
The dataset contains information from the last 50 years of Scooby Doo episodes and movies.  
I wanted to see how the motives changed for the monsters in Scooby Doo over the decades using treemaps.  
I've never tried to animate any of my chart attempts, so I also gave that a go!  
_Tidy Tuesday is a weekly data project for R. Find out more [here](https://github.com/rfordatascience/tidytuesday)._

## What where the monsters' motives in Scooby Doo over the decades?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(stringr)
library(tidyverse)
library(tidytuesdayR)
library(ggplot2)
library(dplyr)
library(lubridate)
library(treemap)
library(gganimate)
library(gapminder)
library(treemapify)
library(gifski)
library(png)

scoobydoo <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-13/scoobydoo.csv')

scoobydoo$year <- year(scoobydoo$date_aired)

```


```{r motives, echo=FALSE, message=FALSE}

total_motives <- scoobydoo %>%
  select(year, motive) %>%
  mutate(year = as.numeric(year)) %>%
  filter(motive != "NULL") %>%
  group_by(year, motive) %>%
  summarize( count = n()) 
  

motive_1960s <- scoobydoo %>%
  select(year, motive) %>%
  filter(str_detect (year, "196")) %>%
filter(motive != "NULL") %>%
  arrange(motive, year) %>%
  group_by(motive) %>%
  summarise(total = n()) %>%
  arrange(desc(total)) 

motive_1970s <- scoobydoo %>%
  select(year, motive) %>%
  filter(str_detect (year, "197")) %>%
filter(motive != "NULL") %>%
  arrange(motive, year) %>%
  group_by(motive) %>%
  summarise(total = n()) %>%
  arrange(desc(total))

motive_1980s <- scoobydoo %>%
  select(year, motive) %>%
  filter(str_detect (year, "198")) %>%
filter(motive != "NULL") %>%
  arrange(motive, year) %>%
  group_by(motive) %>%
  summarise(total = n()) %>%
  arrange(desc(total))

motive_1990s <- scoobydoo %>%
  select(year, motive) %>%
  filter(str_detect (year, "199")) %>%
filter(motive != "NULL") %>%
  arrange(motive, year) %>%
  group_by(motive) %>%
  summarise(total = n()) %>%
  arrange(desc(total))

motive_2000s <- scoobydoo %>%
  select(year, motive) %>%
  filter(str_detect (year, "200")) %>%
filter(motive != "NULL") %>%
  arrange(motive, year) %>%
  group_by(motive) %>%
  summarise(total = n()) %>%
  arrange(desc(total))

motive_2010s <- scoobydoo %>%
  select(year, motive) %>%
  filter(str_detect (year, "201")) %>%
filter(motive != "NULL") %>%
  arrange(motive, year) %>%
  group_by(motive) %>%
  summarise(total = n()) %>%
  arrange(desc(total))

motive_2020s <- scoobydoo %>%
  select(year, motive) %>%
  filter(str_detect (year, "202")) %>%
filter(motive != "NULL") %>%
  arrange(motive, year) %>%
  group_by(motive) %>%
  summarise(total = n()) %>%
  arrange(desc(total))

```

``` {r motive treemap, echo=FALSE}

treemap(motive_1960s, 
        index="motive",
        vSize="total",
        type="index",
        title="Monster motives in the 1960s")

treemap(motive_1970s, 
        index="motive",
        vSize="total",
        type="index",
        title="Monster motives in the 1970s")

treemap(motive_1980s, 
        index="motive",
        vSize="total",
        type="index",
        title="Monster motives in the 1980s")

treemap(motive_1990s, 
        index="motive",
        vSize="total",
        type="index",
        title="Monster motives in the 1990s")

treemap(motive_2000s, 
        index="motive",
        vSize="total",
        type="index",
        title="Monster motives in the 2000s")

treemap(motive_2010s, 
        index="motive",
        vSize="total",
        type="index",
        title="Monster motives in the 2010s")

```

``` {r animate treemap, include=FALSE}

p <- ggplot(total_motives, aes (
            label = motive,
            area=count,
            fill=motive)) +
  geom_treemap(layout = "fixed") +
  theme(legend.position="none") +
  geom_treemap_text(layout = "fixed", place = "centre", grow = TRUE, colour = "white") +
  transition_manual(as.integer(year)) +
   ease_aes('linear') +
  labs(title = "Year: {current_frame}")


anim_save("scoobymotives2.gif", p, fps = 2, nframes = 44)

```
