---
title: "Tidy Tuesday - Olympics"
author: "Amy Mitchell-Whittington"
date: "27/07/2021"
output: github_document
---

```{r setup, include=FALSE}
library(stringr)
library(tidyverse)
library(tidytuesdayR)
library(ggplot2)
library(dplyr)
library(lubridate)
library(convertr)
library(dplyr)
library(RColorBrewer)
library(gganimate)
library(tweenr)
library(forcats)
library(scales)

olympics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-27/olympics.csv')

medals <- olympics %>%
  filter(medal != "NA")

head(olympics)
```

156 medals 
3578 competitions
only in summer Olympics
seven Games(minus 1916, 1940, 1944)
1916 (WW) 1940 & 1944 (WW)


which countries won the most medals in art categories over the years?
-use a graph that animates for which countries won most gold, silver, bronze each year.

``` {r data clean, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)

# look at the arts categories

art_comp <- olympics %>%
  select(year, sport, event, sex, medal,team) %>%
  filter(sport == "Art Competitions",
         medal != "NA") %>%
  mutate_at("event", str_replace, "Art Competitions Mixed", "") %>% #to remove undesired characters when they're constant.
  arrange(team)

art_country <- art_comp %>%
  select(year, medal, team) %>%
  group_by(year, medal, team) %>%
  dplyr::summarize(total_medals = n()) %>% # to calculate total medals by year and team 
  ungroup() %>%
  arrange(-total_medals) 

art_country$medal <- factor(art_country$medal, levels = c("Gold", "Silver", "Bronze"))

```


``` {r art plot, include=FALSE}

l <-  ggplot(art_country) +
  geom_linerange(aes(
              x= team, 
              ymin = 0, ymax = total_medals, colour=medal),
              position = position_dodge(width=1)) +
  scale_y_continuous(limits=c(0,6)) +
  geom_point(aes ( x =team, y = total_medals, colour = medal),
             position = position_dodge(width = 1)) +
  scale_color_manual(values = c("#cc9933", "#999999", "#996600")) +
  coord_flip() +
  facet_wrap(~year, ncol=4, scales = "free_y") +
  labs(
    title = "Medals won by country in the Olympic Arts Competitions from 1912-1948",
    y=NULL,
    x=NULL,
    legend = "Type of medals", 
    caption = "Data: Kaggle/TidyTuesday."
  ) + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) 

```

```{r tug data clean, include=TRUE}

# tug was showing as a tibble, not a dataframe, so assigning it recycled values (gold=3, silver=2, bronze=1) for the medals to work in a graph format didn't work. 
#have to create a non-tibble version to recycle the value.


tug <- olympics %>%
  select(year, sport, medal, team) %>%
  filter(sport=="Tug-Of-War", medal != "NA") %>%
group_by(medal, team) %>%
  unique()

tug_df <- as.data.frame(tug)


# tug was showing as a tibble, not a dataframe, so assigning it recycled values (gold=3, silver=2, bronze=1) for the medals to work in a graph format didn't work. 
#have to create a non-tibble version to recycle the value.

tug_df <- as.data.frame(tug)

#tug_df <- tug_df %>%
 # mutate(medal_count=case_when(
  #  .$medal=="Gold" ~ 3,
   # .$medal=="Silver" ~ 2,
    #.$medal=="Bronze" ~ 1))

#to change the order of the medals

tug_df$medal <- factor(tug_df$medal, levels = c("Gold", "Silver", "Bronze"))

tug_1908 <- tug %>%
  filter(year == "1908") %>%
  dplyr::summarize(total_medals = n()) 


```

```{r plot}
knitr::opts_chunk$set(echo=TRUE)


tug_df %>% 
  ggplot(aes(x = team, y = medal)) +
  geom_point(tug_df %>% 
               filter(medal == "Gold"), 
              mapping = aes( y = 1), 
             size = 15, 
             color = "gold", 
             show.legend = FALSE) +
 geom_text(tug_df %>% 
               filter(medal == "Gold"),
             mapping = aes(label = year, y = 1)) +
  geom_point(tug_df %>% 
               filter(medal == "Silver"), 
             mapping = aes(y = 2), 
             size = 15, color = "#d6d6d6", 
             show.legend = FALSE) +
  geom_text(tug_df %>% 
               filter(medal == "Silver"),
            mapping = aes(label = year, y = 2)) +
  geom_point(tug_df %>% 
               filter(medal == "Bronze"), 
              mapping = aes(y = 3), 
             size = 15, 
             color = "#cd7f32", 
             show.legend = FALSE) +
  geom_text(tug_df %>% 
               filter(medal == "Bronze"),
            mapping = aes(label = year, y = 3)) +
  coord_flip() +
  labs(
    title = "Medals won by participants in the Tug-Of-War Competitions between 1912-1948",
    y=NULL,
    x=NULL,
    legend = "Type of medals", 
    caption = "Data: Kaggle/TidyTuesday."
  ) + 
  theme_minimal() +
  theme(panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_blank(),
        axis.text.x = element_blank()) +
  theme(plot.title = element_text(hjust = 0.5))

```







plot <- ggplot(tug_df) +
  geom_point(aes ( x =team, y = medal_count, colour = medal),
             size = 10,
             position = position_dodge(width = 1)) +
  scale_color_manual(values = c("#cc9933", "#999999", "#996600")) +
  geom_linerange(aes(
              x= team, 
              ymin = 0, ymax = medal_count, colour=medal),
              position = position_dodge(width=1)) +
  coord_flip() +
  labs(
    title = "Medal winners in the Tug-Of-War Competitions between 1912-1948",
    y=NULL,
    x=NULL,
    legend = "Type of medals", 
    caption = "Data: Kaggle/TidyTuesday."
  ) + 
  theme_minimal() +
  theme(panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_blank(),
        axis.text.x = element_blank()) +
  theme(plot.title = element_text(hjust = 0.5)) 


show(plot)


geom_linerange(aes(
              x= team, 
              ymin = 0, ymax = medal_count, colour=medal),
              position = position_dodge(width=1)) +
  scale_y_continuous(limits=c(0,3)) +
```
