---
title: "Tidy Tuesday - U.S. drought"
author: "Amy Mitchell-Whittington"
date: "21/07/2021"
output: github_document
---

## U.S. Drought

This week's data comes from the [U.S. Drought Monitor](https://droughtmonitor.unl.edu/DmData/DataDownload/ComprehensiveStatistics.aspx).

The dataset details the drought level across U.S. states from 2001-2021. I wanted to look at the number of people who have experienced severe, extreme, and/or exceptional drought conditions over this time in the western states of the U.S.

Data Reference: The U.S. Drought Monitor is jointly produced by the National Drought Mitigation Center at the University of Nebraska-Lincoln, the United States Department of Agriculture, and the National Oceanic and Atmospheric Administration. Map courtesy of NDMC.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(stringr)
library(tidyverse)
library(tidytuesdayR)
library(ggplot2)
library(dplyr)
library(lubridate)
library(convertr)
library(scales)
library(tidyselect)
library(plotly)
library(plyr)

drought <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-20/drought.csv')

# to prevent scientific notation in graph:
options(scipen=999)
```


```{r data clean, echo=FALSE, message=FALSE}

# remove Puerto Rico 

drought_states <- drought %>%
  filter(state_abb != "PR") 

# remove null figures 

drought_states[drought_states == 0] <- NA

# need to separate the western states of U.S. and remove none, D0, D1 variables

western_states_drought <- drought_states %>%
  filter(state_abb %in% c("AK", "AZ", "CA", "CO", "ID", "MT", "NV", "NM", "OR", "UT", "WA", "WY")) %>%
  filter(drought_lvl %in% c("D2", "D3", "D4")) %>%
  na.omit()

# need to group sum pop_total by two variables 
  
western_states <- western_states_drought %>%
  select(state_abb, valid_start, drought_lvl, pop_total) %>%
  ddply(.(drought_lvl, valid_start), summarize, total = sum(pop_total)) 
```

```{r plot, echo=FALSE, message=FALSE}

COLORS = c("#F08000", "#D9490B", "#A81308")

wss <- western_states %>%
  ggplot +
  geom_area(aes(valid_start, total, fill=drought_lvl)) +
  scale_fill_manual(labels = c(
    "severe drought", "extreme drought", "exceptional drought"),
    values = COLORS) +
  scale_y_continuous( 
    breaks = seq(0, 600000000, 10000000),
    labels = label_number(scale = 1/1000000)) +
  labs( 
    title="Drought across the decades",
    subtitle="Total number of people in U.S. western states who experienced \n severe, extreme, or exceptional drought over the past two decades.",
    caption="data: The U.S. Drought Monitor.", 
    y="Number of people (million)",
    x="Year",
    fill="Type of drought") +
  theme(
    plot.caption = element_text(size = 8),
    plot.title = element_text(size = 14, hjust=0.5),
    plot.subtitle = element_text(hjust=0.5, size=10)
  )

ggplotly(wss)
```
